---
title: "Thesis_Public_Charging"
author: "Sarah Newton"
date: "2024-02-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r env_setup, include=FALSE}
## Base packages
library(tidyverse)
library(jsonlite)
library(tmaptools)
library(gtsummary)
library(stargazer)
library(stringr)
library(wesanderson)
library(knitr)
library(rlist)

## GiS handling
library(osmdata)
library(sf)
library(googleway)
library(secret)
library(ggmap)

## Routing Analysis
library(sfnetworks)
library(igraph)
library(netdiffuseR)
library(tidygraph)
library(cppRouting)
library(tidygraph)
library(RANN)
library(dodgr)
library(stplanr)

# Accessibility
library(accessibility)
library(r5r)
# Network cleaning
library(rgrass)
library(link2GI)
#remotes::install_github('thomasp85/tidygraph')
#https://r-spatial.org/r/2019/09/26/spatial-networks.html
#PCA and Correlation
library('corrr')
library(ggcorrplot)
library(factoextra)

### Enviro Setup
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path))

opts_chunk$set(eval=FALSE,
               include=FALSE,
               echo = FALSE, 
               message = FALSE, 
               warning = FALSE, 
               class.source = "bg-primary")


#### Google Places API key
key <- paste(read_file(paste0(dirname(getwd()), "/gm_api_key.txt")), collapse="\n")
```

```{r load_environment, eval=TRUE}
## Placeholders to be able to save and load heavy datasets and variables
#load('Charge_Environment.RData')
#save.image(file='Charge_Environment.RData')
#save(streets, file="streets.RData")
#save(distances_per_point, file="distances_per_point.RData")
```

## GiS setup and coordinations

```{r gis_setup, include=FALSE}
### Gothenburg subdivisions data

# DeSO subdivisions data from .gpkg
geo <- st_read(dsn="data/DeSO_2018_v2.gpkg")
# ReGSO subdivisioms data from .gpkg
regSO_deso_table <- read_csv2("data/kopplingstabell-deso-regso-20210702.csv")

regSO_deso_table <- regSO_deso_table %>% 
  filter(Kommunnamn == "Göteborg") %>%
  rename(deso = DeSO, regso = RegSO) %>%
  select(deso, regso)

# Create GOT kommon polygon from .gpkg file DeSOs
gothenburg.subdiv <- geo %>% filter(kommunnamn == "Göteborg") %>% 
# Filter out category A and B deso areas
# https://www.scb.se/hitta-statistik/regional-statistik-och-kartor/regionala-indelningar/deso---demografiska-statistikomraden/
                             filter(grepl("C", deso) == TRUE)

# Convert Gothenburg subdivision polygons to same crs 4326 as osm data
gothenburg.subdiv <- gothenburg.subdiv %>% st_transform(., 4326)
### OSM data setup
#first option (not recommended in this case)
#town <- 'Gothenburg' 
#location <- town %>% opq()

# Get GOT boundary using coodinates
coords <- st_bbox(gothenburg.subdiv)
#coords <- matrix(c(11.6, 12.05, 57.63, 57.80), byrow = TRUE, nrow = 2, ncol = 2, dimnames = list(c('x','y'),c('min','max'))) 
location <- opq(st_bbox(gothenburg.subdiv))

# Alternative approach for getting GOT boundary
# Get the boundary for Gothenburg
# got.boundary <- opq(bbox = "Gothenburg, Sweden") %>%
#   add_osm_feature(key = 'admin_level', value = '7') %>% 
#   osmdata_sf %>% unique_osmdata

# got.region <- got.boundary$osm_multipolygons %>% filter(osm_id == '935611')
# 
# neg_buffer <- st_buffer(got.region, -100) # in meters
# got_municipalities_poly <- got.boundary$osm_multipolygons[neg_buffer, ]
```


```{r p2p_charge_point_cleaning, include=FALSE}
## Scraped data from "Place to Plug" website
## Scraped from Placetoplug.com 31-01-2024
## Address: https://placetoplug.com/en/charging-stations/Sweden/Gothenburg/1
## Search term Gothenburg, Sweden
## 380 results, 38 pages.

json <- jsonlite::fromJSON(txt="data/Charging_Points-place-to-plug.json")
p2p.data <- json$chargingZones

p2p.data <- p2p.data %>% select(id, name, address, isFastCharge, powerRange, plugTypes, status, vehicleTypes, -`__typename`)

p2p.data <- p2p.data %>% mutate(full_address = gsub(" , ", "", paste(ifelse(is.na(address$street), name, address$street), 
                                         address$city, 
                                         address$subdivision)))
getDetailsUrl <- function (x) {
  return(paste0("https://placetoplug.com/en/charging-stations/Sweden/", 
                x$full_address, "/", 
                x$id, "/", 
                str_replace(x$name, " ", "_")))
  
}
p2p.data <- p2p.data %>% mutate(url = getDetailsUrl(.))

# Example of charge point demographics that are available
ex.chargepoint.demographics <- fromJSON("data/test_charge_details.json")
#view(ex.chargepoint.demographics$data$chargingZone$stations %>% as_tibble())

# Set up google api key for ggmap on first use
# register_google(key = key, write=TRUE)
# Fetch long and lat for the charging point
# p2p.long_lat <- geocode(tolower(p2p.data$full_address))
p2p.long_lat <- read_csv("data/p2p_data_long_lat.csv")
p2p.data <- p2p.data %>% mutate(p2p.long_lat)
# Plug types
#IEC_62196_T2
#IEC_62196_T2_COMBO
#CHADEMO
#OTHER
#DOMESTIC_F
#TESLA_S_EU

# Clean data
p2p.data <- p2p.data %>% mutate(vehicleTypes.car = ifelse(grepl("CAR", p2p.data$vehicleTypes), TRUE, FALSE),
                                vehicleTypes.bike = ifelse(grepl("BICYCLE", p2p.data$vehicleTypes), TRUE, FALSE),
                                vehicleTypes.motorbike = ifelse(grepl("BIKE", p2p.data$vehicleTypes), TRUE, FALSE),
                                status = factor(status),
                                powerRange.min = as.numeric(p2p.data$powerRange$min),
                                powerRange.max = as.numeric(p2p.data$powerRange$max))

p2p.data <- p2p.data %>% na.omit(powerRange.min, powerRange.max) %>% select(id, full_address, lon, lat, status,
                                             vehicleTypes.car, vehicleTypes.bike, vehicleTypes.motorbike, 
                                             isFastCharge, powerRange.min, powerRange.max
                                              )
p2p.data <- p2p.data %>% mutate(lng = lon) %>% filter(powerRange.min > 0)
# Filter out charge points outside of Gothenburg coords
p2p.data <- p2p.data %>% filter(between(lng, coords[1], coords[3]) &
                                                    between(lat,coords[2], coords[4]))

charge_points.p2p <- p2p.data %>% select(lat, lng)

# spatial join of subdivisions and charge points
charge_points.p2p.geo <- st_as_sf(charge_points.p2p, 
                                  coords = c("lng", "lat"), crs = 4326) %>% rownames_to_column(., "charge_id")

# Convert Gothenburg subdivision polygons to same crs 4326 as osm data
gothenburg.subdiv <- gothenburg.subdiv %>% mutate(area = st_area(.)) %>%
  st_as_sf(., coords = c("x","y"), crs = 4326)

charge_points.p2p.geo <- st_filter(charge_points.p2p.geo, gothenburg.subdiv)

# Save cleaned and merged data as csv
# write_csv(p2p.long_lat, "p2p_data_long_lat.csv")
# write_csv(p2p.data, "p2p_data_complete.csv")
# p2p.data <- read_csv("p2p_data_complete.csv")

```


```{r google_setup_option, include=FALSE}

# Google places API alternative using 'googleway'
# res <- google_places(
#   search_string = NULL,
#   location = c(57.708870, 11.974560),
#   radius = 5000,
#   rankby = NULL,
#   keyword = NULL,
#   language = NULL,
#   name = NULL,
#   place_type = "electric_vehicle_charging_station ",
#   price_range = NULL,
#   open_now = NULL,
#   page_token = NULL,
#   simplify = TRUE,
#   curl_proxy = NULL,
#   key = key,
#   radar = NULL
# )
# colnames(res$results[1,])
# charge_points <- res$results$geometry$location
# colnames(charge_points)
# colnames(charge_points.p2p)
```

```{r fetch_road_network_osm, include=FALSE}

# Set the relevant tags and values for each street type needed
main_st <- data.frame(type = c("motorway","trunk","primary","motorway_junction","trunk_link","primary_link","motorway_link"))
st <- data.frame(type = available_tags('highway')$Value)
non_vehicle <- data.frame(type = c("footway", "path", "steps", "cycleway", "corridor", "pedestrian", "track", "bridleway", "bus_guideway", "busway", "bus_stop", "construction", "elevator", "proposed", "street_lamp"))

# Subset by those not in main and non_vehicle streets
st <- subset(st, !type %in% main_st$type & !type %in% non_vehicle$type)

## Save final list as character strings
st <- as.character(st$type)
main_st <- as.character(main_st$type)

main_streets <- location %>%
  add_osm_feature(key = "highway", 
                  value = main_st) %>% osmdata_sf()

streets <- location %>%
  add_osm_feature(key = "highway", 
                  value = st) %>% osmdata_sf()

### Extra map objects if needed for visualising
# water <- location %>%
#   add_osm_feature(key = "natural", 
#                   value = c("water", "lake", "bay", "strait")) %>% osmdata_sf()
# parks <- location %>% add_osm_feature(key = "leisure", 
#                                       value = c("park","nature_reserve","recreation_ground","golf_course","pitch","garden")) %>%
#                       osmdata_sf()
# forest <- location %>% add_osm_feature(key = "landuse", value = c("forest")) %>%
#  osmdata_sf()
# 
# land <- location %>%
#   add_osm_feature(key = "natural", 
#                   value = available_tags('natural')$Value) %>% osmdata_sf()

```

## Methods

```{r Clean_network, include=FALSE}

# Add data to GRASS spatial database 
# write_VECT(
#   x = streets$osm_lines, 
#   vname = 'gothenburg', 
#   flags = 'overwrite'
# )

# Execute the v.clean tool
# rgrass::execGRASS("g.proj", flags = c("c", "quiet"), proj4 = proj4)
# execGRASS(
#   cmd = 'v.clean', 
#   input = 'muenster_center', 
#   output = 'muenster_cleaned',        
#   tool = 'break', 
#   flags = c('overwrite', 'c')
# )
# 
# # Read back into R
# use_sf()
# muenster_center <- readVECT('muenster_cleaned') %>%
#   rename(geometry = geom) %>%
#   select(-cat)
```

### Create network graph

```{r setup_network_graph, include=FALSE}
##### Preparing Graph of nodes and edges #####

### Bind the relevant streets and paths into one set 
got_streets <- rbind((streets$osm_lines %>% select(highway)), 
                     main_streets$osm_lines %>% select(highway))

# Give each edge a unique index
edges <- got_streets %>% 
  select(highway) %>%
  mutate(edgeID = c(1:n()))

# Create nodes at the start and end point of each edge
nodes <- edges %>%
  st_coordinates() %>%
  as_tibble() %>%
  rename(edgeID = L1) %>%
  group_by(edgeID) %>%
  slice(c(1, n())) %>%
  ungroup() %>%
  mutate(start_end = rep(c('start', 'end'), times = n()/2))

# Give each node a unique index
nodes <- nodes %>%
  mutate(xy = paste(.$X, .$Y)) %>% 
  mutate(nodeID = group_indices(., factor(xy, levels = unique(xy)))) %>%
  select(-xy)

# Combine the node indices with the edges
source_nodes <- nodes %>%
  filter(start_end == 'start') %>%
  pull(nodeID)

target_nodes <- nodes %>%
  filter(start_end == 'end') %>%
  pull(nodeID)

# Merge source nodes and target nodes
edges <- edges %>%
  mutate(from = source_nodes, to = target_nodes)

# Remove duplicate nodes
nodes <- nodes %>%
  distinct(nodeID, .keep_all = TRUE) %>%
  select(-c(edgeID, start_end)) %>%
  st_as_sf(coords = c('X', 'Y')) %>%
  st_set_crs(st_crs(edges))

# Create a graph from the adjacency matrix
graph <- tbl_graph(nodes = nodes, edges = as_tibble(edges), directed = FALSE)

# Identify isolate nodes in the graph
# isolated(as_adjacency_matrix(graph)) %>% as_tibble() %>% filter(value==TRUE)

# The activate() verb specifies if we want to manipulate the edges or the nodes. 
graph <- graph %>%
  activate(edges) %>%
  filter(!edge_is_multiple()) %>%
  filter(!edge_is_loop()) %>%
  mutate(length = st_length(geometry))

# Subdivide the edges
graph <- convert(graph, to_spatial_subdivision)
# Smooth
# Smoothing reduces precision so will not use
# graph <- convert(graph, to_spatial_smooth, summarise_attributes = list(length = "sum"))

graph %>% activate(edges) %>% as_tibble() %>% filter(is.na(length))# %>% st_as_sf()
# graph <- graph %>% activate(edges) %>% as_tibble() %>% filter(!is.na(length)) %>% st_as_sf()

```

```{r setup_network_graph}
##### Preparing the road network graph of nodes and edges #####

### Bind the relevant streets and paths into one set 
got_streets <- rbind((streets$osm_lines %>% select(highway)), 
                     main_streets$osm_lines %>% select(highway))

got_streets_net <- as_sfnetwork(got_streets, directed = FALSE)

# Check the CRS to ensure it's correct before continuing: EPSG:4326
st_crs(got_streets_net)

# Use the “spatial morpher” function to create a subdivision of edges
got_streets_net <- convert(got_streets_net, to_spatial_subdivision)

# removal of nodes that have only two edges connected, also called “smoothing of pseudo-nodes”
# got_streets_simple <- convert(got_streets_net, to_spatial_smooth)
# !! For this case simplification may be counter productive due to the need to measure isodistances and reduces precision

# Add length of edges as a new variable. Remove looping edges
got_streets_net <- got_streets_net %>% 
  activate("edges") %>% 
  filter(!edge_is_multiple()) %>%
  filter(!edge_is_loop()) %>%
  mutate(length = edge_length())

# Remove "islands" of nodes and edges isolated from the wider road network. 
# Threshold is set to 6
got_streets_net <- got_streets_net %>% 
  activate(nodes) %>% 
  mutate(neighbourhood = local_size(order = 6)) %>% 
  filter(neighbourhood > 5)

# Add id" to each nodes in the network based on rowname index
# !!Make sure that this is the last step done to the network otherwise filtering and other mutations may impact the correctness of this.
got_streets_net <- got_streets_net %>% 
  activate("nodes") %>% 
  mutate(id = c(1:n()))


```


```{r nodes_to_coords_matrix}
### Convert nodes to coordinates matrix
got_streets.df <- got_streets_net %>%
  activate(nodes) %>%
  as_tibble()

got_streets.sf <- got_streets.df %>% st_as_sf()

### Convert nodes to coordinates matrix
got_streets_coords <- got_streets_net %>%
  activate(nodes) %>%
  as_tibble() %>% st_as_sf() %>%
  st_coordinates()

### Convert charging points to coordinates matrix
charge_points.long_lat.matrix <- as.matrix(st_coordinates(charge_points.p2p.geo)) %>% na.omit()
```

### KNN 5 "as the crow flies"

Calculate the closest charge point to a node/intersection using K nearest neighbour (5). This method is the simplest measure for calculating distance, and does NOT take into consideration the surrounding road network

```{r setup_knn_nearest_point}

#### FINDING NEAREST POINTS #####

### As the crow flies method
### RANN nn2
### Find the K nearest neighbour charging point to each node (KNN)
knn.5 <- nn2(charge_points.long_lat.matrix, 
    query = got_streets_coords,
    k = 5,
    treetype = "kd",
    eps = 0)

#### FINDING NEAREST NODES #####

# Calculate closest nodes to each charge point using KNN 5
knn.5.nodes <- nn2(got_streets_coords, 
             query = charge_points.long_lat.matrix,
             k = 5,
             treetype = "kd",
             eps = 0)

# Calculate closest nodes to each charge point using KNN 5 Radius
knn.radius <- nn2(got_streets_coords, 
             query = charge_points.long_lat.matrix,
             k = 5,
             treetype = "kd", 
             searchtype = "radius", 
             radius = 0.5,
             eps = 0)
```

```{r extract_closest_from_knn_5}
### Function for selecting the node with the closest distance from KNN
getMinKnn <- function(i, ids, distances) {
  min_dist_i <- which.min(distances[i,])
  return(data.frame(ids[i,min_dist_i], distances[i,min_dist_i]))
}

closest_charge <- data.frame()
for (i in 1:nrow(knn.5$nn.idx)) {
  #print(paste("running:", i,  "of", as.character(nrow(knn.5$nn.idx))))
  new_elements <- getMinKnn(i, knn.5$nn.idx, knn.5$nn.dists)
  closest_charge <- rbind(closest_charge, new_elements)
}

colnames(closest_charge) <- c("charge_id", "distance")
got_streets_net.knn <- cbind(got_streets_coords, closest_charge) %>% 
  mutate(Lon = X, Lat = Y) %>% select(-X, -Y)
```

```{r extract_closest_from_knn_5.nodes}
### Function for selecting the node with the closest distance from KNN
closest.nodes <- data.frame()
for (i in 1:nrow(knn.5.nodes[[2]])) {
  new_elements <- getMinKnn(i, knn.5.nodes$nn.idx, knn.5.nodes$nn.dists)
  closest.nodes <- rbind(closest.nodes, new_elements)
}

colnames(closest.nodes) <- c("node_id", "distance")
```

```{r extract_closest_from_knn_5.radius.nodes}
### Function for selecting the node with the closest distance from KNN. 
### This will be used to set the snap point for the charge point so that it is directly on the network vertices
closest.nodes.radius <- data.frame()
for (i in 1:nrow(knn.radius[[2]])) {
  new_elements <- getMinKnn(i, knn.radius$nn.idx, knn.radius$nn.dists)
  closest.nodes.radius <- rbind(closest.nodes.radius, new_elements)
}

colnames(closest.nodes.radius) <- c("node_id", "distance")
```

### Snap charge points to nodes

"A snap distance tolerance is used to control where snapping is performed. Snapping one geometry to another can improve robustness for overlay operations by eliminating nearly-coincident edges (which cause problems during noding and intersection calculation). Too much snapping can result in invalid topology being created, so the number and location of snapped vertices is decided using heuristics to determine when it is safe to snap. This can result in some potential snaps being omitted, however."

```{r snap_points_to_nodes, include=FALSE}
### Snap charging points to nodes
snap_matrix <- closest.nodes %>%
  rownames_to_column(., var = "charge_id") %>%
  #arrange(distance) %>% 
  #distinct(charge_id, .keep_all = TRUE) %>%
  #arrange(charge_id) %>%
  select(node_id, charge_id)

# Based on the closest network node to a charge point, snap point
getSnapPoint <- function(node) {
  i <- match(node, snap_matrix[,2])
  snap_node <- as.integer(snap_matrix[i,]$node_id)
  snap_point <- got_streets.df[snap_node,]$geometry
  return(snap_point)
}

# Check to ensure working function and no NAs
# na_charge_list <- data.frame()
# checkNa <- function(charge) {
#   i <- match(charge, snap_matrix[,2])
#   snap_node <- as.integer(snap_matrix[i,]$node_id)
#   snap_point_id <- got_streets.sf[snap_node,]$nodeID
#   return(data.frame(snap_node, charge))
# }
# 
# for (i in 1:nrow(charge_points.p2p.geo)) {
#   new_elements <- checkNa(charge_points.p2p.geo$charge_id[i])
#  # To check which values are null
#   na_charge_list <- ifelse(is.na(new_elements[1]), 
#        rbind(na_charge_list, new_elements[2]), na_charge_list)
# }
# na_charge_list <- na_charge_list[[1]]

getNodePoint <- function(charge) {
  i <- match(charge, snap_matrix[,2])
  # Get nodeID of the point to snap to
  snap_node <- as.integer(snap_matrix[i,]$node_id)
  #snap_point_id <- got_streets.sf[snap_node,]$nodeID
  #print(paste(charge, i, snap_node, snap_point_id))
  return(snap_node)
}

charge_points.p2p.geo.nodes <- charge_points.p2p.geo %>% 
  mutate(nodeId = mapply(getNodePoint, charge_id)) %>%
  select(charge_id, nodeId) %>%
  as_tibble() %>%
  st_as_sf(crs = 4326)

charge_points.p2p.geo.nodes %>% filter(is.na(nodeId))

charge_points.p2p.geo.snaps <- charge_points.p2p.geo %>% 
  mutate(geometry = mapply(getSnapPoint, charge_id)) %>%
  select(charge_id, geometry) %>%
  as_tibble() %>%
  st_as_sf(crs = 4326)

## Merge the charge point data with the original graph
### TODO
# Get the number of chargers per point here, rather than defaulting to 1
got_streets_net.merged <- got_streets_net %>% 
  activate(nodes) %>% 
  mutate(charge_points = ifelse(id %in% charge_points.p2p.geo.nodes$id, 1, 0))# %>% drop_na()

```

```{r plot_network_graph, eval=TRUE, include=TRUE}
ggplot() +
  geom_sf(data = got_streets_net %>% activate(edges) %>% as_tibble() %>% st_as_sf(), linewidth = 0.2, color="#1e00be") + 
  geom_sf(data = got_streets_net %>% activate(nodes) %>% as_tibble() %>% st_as_sf(), size=0.1, color="#1e00be") +
  geom_sf(data=gothenburg.subdiv, alpha=0.4, linewidth=0.5, colour="#8778D7", fill="#A753AF") +
  coord_sf(lims_method = "geometry_bbox",
           xlim = c(11.71, 12.16), 
           ylim = c(57.58, 57.86), expand = TRUE, clip = "on") +
  geom_sf(data = charge_points.p2p.geo, size=0.9, color="#90C59B") +
  geom_sf(data = charge_points.p2p.geo.snaps, size=0.9, color="red") +
  coord_sf(#lims_method = "geometry_bbox",
    xlim = c(11.71, 12.16), 
    ylim = c(57.58, 57.86), expand = TRUE, clip = "on") +
  labs(x="", y="") +
  theme_minimal() + theme(legend.position = "")
```

```{r filter_node_by_charge_points}
# nodes.charge_only <- edges %>%
#   st_coordinates() %>%
#   as_tibble() %>%
#   rename(edgeID = L1) %>%
#   group_by(edgeID) %>%
#   slice(c(1, n())) %>%
#   ungroup() %>%
#   mutate(start_end = rep(c('start', 'end'), times = n()/2))
# 
# # Give each node a unique index
# nodes.charge_only <- nodes.charge_only %>%
#   mutate(xy = paste(.$X, .$Y)) %>% 
#   mutate(nodeID = group_indices(., factor(xy, levels = unique(xy)))) %>%
#   select(-xy)
# 
# nodes.charge_only <- nodes.charge_only %>%
#   distinct(nodeID, .keep_all = TRUE) %>%
#   filter(nodeID %in% snap_matrix$node_id)
# 
# edges.charge_only <- edges %>% 
#   filter(edgeID %in% nodes.charge_only$edgeID)
# 
# nodes.charge_only <- nodes.charge_only %>%
#   select(-c(edgeID, start_end)) %>%
#   st_as_sf(coords = c('X', 'Y')) %>%
#   st_set_crs(st_crs(edges))
# 
# got_streets_net.charge_only <- got_streets_net %>% filter(edgeID %in% edges.charge_only$edgeID)

```

### Dijkstra's graph search (shorted path)

Alternative more "accurate" measure for shortest paths what considering road networks and driveability.

```{r st_planr_dodgr_cleaning}
#dodgr::igraph_to_dodgr(got_streets_net)
```

```{r loop_dijkstras_charge_points}
# https://github.com/ropensci/stplanr/issues/261 Talking about issues with stplanr
# Loop over each charge point node and find the shortest path from a charge point to a nearby node
cp.test <- charge_points.p2p.geo.nodes %>% na.omit()

test <- data_frame()
for(i in 1:nrow(cp.test)) {
  # select closest nodes
  closest_nodes <- knn.5.nodes$nn.idx[i,] %>% as_tibble() %>% filter(value != cp.test[i,]$nodeId)
  print(paste("Start:", cp.test[i,]$charge_id))
  # loop to get shortest path for each
  for(ii in 1:nrow(closest_nodes)) {
    from_cp_node <- got_streets_net.merged %>%
                    activate(nodes) %>%
                    filter(id == cp.test[i,]$nodeId) %>%
                    pull(id)
    
    to_node <- got_streets_net.merged %>%
                    activate(nodes) %>%
                    filter(id == as.character(closest_nodes[ii,])) %>%
                    pull(id)
      
    # print(paste("from node:", from_cp_node, "to node:",  to_node))
      path <- shortest_paths(
              graph = got_streets_net.merged,
              from = from_cp_node,
              to = to_node,
              #mode = "all",
              output="both",
              algorithm ="dijkstra", # Ensure to use Dijkstra's algorithm
              weights = got_streets_net.merged %>% activate(edges) %>% pull(length)
            )
      
    path_graph <- got_streets_net.merged %>%
                  subgraph.edges(eids = path$epath %>% unlist()) %>%
                  as_tbl_graph()
    
   distance_data <- path_graph %>%
          activate(edges) %>%
          as_tibble() %>%
          summarise(length = sum(length))
   
    print(paste("Distance:", distance_data))
    new_distances <- data_frame(data_frame(from_cp_node, to_node, distance_data))
    test <- rbind(test, new_distances)
  }
  
  print("")
}

dijkstras_cp_to_node <- test
colnames(dijkstras_cp_to_node) <- c("from_node", "to_node", "distance")
view(dijkstras_cp_to_node.avg)

# Group by the change point and get average distance to surrounding nodes
dijkstras_cp_to_node.avg <- dijkstras_cp_to_node %>% 
  group_by(from_node) %>% 
  summarise(avg_travel_dist = mean(distance)) %>% 
  arrange(avg_travel_dist)
```

```{r loop_dijkstras_nodes}
## Loop over each node in the entire road network and find the shortest path to a charge point
gothenburg.subdiv_net.merged <- st_join(got_streets_net, gothenburg.subdiv, join = st_within, left=TRUE)

#nodes.bergsjön <- gothenburg.subdiv_net.merged %>% filter(deso == '1480C3560') %>% activate(nodes) %>% as_tibble()
nodes.test <- got_streets_net %>% activate(nodes) %>% as_tibble()# %>% filter(id %in% nodes.bergsjön$id) %>% arrange(id)

test.nodes <- data_frame()
for(i in 1:nrow(nodes.test)) {
  # select current node in iteration
  current_node <- nodes.test[i,]$id
  print(paste("Start:", current_node))
  # Get closest charge point
  closest_cp <- cp.test %>% 
    as_tibble() %>% 
    filter(charge_id == closest_charge[current_node,]$charge_id)
  # If not found, skip to next item
  ifelse(nrow(closest_cp) == 0, next, NA)
  print(paste("for charge point:", closest_cp$charge_id))
    # get shortest path for the current node to it's closest charge point
    from_node <- got_streets_net %>%
                    activate(nodes) %>%
                    filter(id == current_node) %>%
                    pull(id) # ! It would be safer to use the vertex rownames here otherwise it can easily be broken if changes are made to the network graph, such as filtering as the shortest_path algorithm does not use the "id" col.
    
    to_node <- got_streets_net %>%
                    activate(nodes) %>%
                    filter(id == closest_cp$nodeId) %>%
                    pull(id)
      
    print(paste("from node:", from_node, "to node:",  to_node))

      path <- shortest_paths(
                graph = got_streets_net,
                from = from_node,
                to = to_node,
                #mode = "all",
                output="both",
                algorithm ="dijkstra", # Ensure to use Dijkstra's algorithm
                weights = got_streets_net %>% activate(edges) %>% pull(length)
              )
    path_graph <-  got_streets_net %>%
                    subgraph.edges(eids = path$epath %>% unlist()) %>%
                    as_tbl_graph()
    
   distance_data <- path_graph %>%
          activate(edges) %>%
          as_tibble() %>%
          summarise(length = sum(length))
   
    print(paste("Distance:", distance_data))
    new_distances <- c(as_tibble(from_node), as_tibble(to_node), distance_data)
    test.nodes <- rbind(test.nodes, new_distances)
  
  print("")
}
```

```{r clean_dijkstras_data}
distances_per_point <- test.nodes
colnames(distances_per_point) <- c("from_node", "to_node", "distance")
view(distances_per_point)
# Save as csv for backup as the function takes 10 hours to run
# write_csv(distances_per_point, "data/dijkstras_per_point.csv")

charge_points.p2p.geo.nodes.dist <- merge(distances_per_point,
                                          charge_points.p2p.geo.nodes, 
                                          by.x="from_node", by.y="nodeId", 
                                          all=TRUE) %>% mutate(charge_id = as.integer(charge_id)) %>%
  arrange(charge_id) %>% st_as_sf()

charge_points.p2p.geo.nodes.dist %>% filter(from_node == to_node)
charge_points.p2p.geo.nodes.dist %>% view()

distances_per_point.sf <- merge(distances_per_point,
                                got_streets_net %>% activate(nodes) %>% as_tibble(), 
                                by.x="from_node", by.y="id", 
                                all=TRUE) %>%
  arrange(from_node) %>% st_as_sf()
# Filter out nodes that are the same as the corresponding charge point as this distance is of course zero.

distances_per_point.sf %>% view()
distances_per_point.sf %>% filter(from_node == to_node)
filtered <- st_filter(distances_per_point.sf, gothenburg.subdiv)

ggplot() +
  # All road network nodes
  #geom_sf(data = got_streets_net %>% activate(nodes) %>% as_tibble() %>% st_as_sf(), size=0.4) +
  # nodes coloured by distance
  geom_sf(data=filtered, size=0.5, aes(color=distance)) +
  # Charge points
  geom_sf(data=charge_points.p2p.geo.nodes.dist, col="blue") +
  # DeSO Subdivisions
  geom_sf(data = gothenburg.subdiv, alpha=0.3, linewidth=0.6, colour="#8778D7", fill="#90C59B") + 
  scale_color_distiller(palette = "Oranges")

joined.nodes <- st_join(distances_per_point.sf, gothenburg.subdiv, join = st_within, left=TRUE)
joined.nodes <- joined.nodes %>% 
  mutate(dijkstras_distance = distance)

avg_dist_deso <- joined.nodes %>% as_tibble() %>% 
  group_by(deso) %>% 
  summarise(nodes = n(),
            #remove NA values from the mean to avoid NA result
            avg_dist = mean(dijkstras_distance, na.rm=TRUE)) %>% 
  arrange(avg_dist) #%>% drop_na()

avg_dist_deso <- merge(avg_dist_deso, regSO_deso_table) %>% 
  arrange(desc(avg_dist))
view(avg_dist_deso)
gothenburg.subdiv.merged <- merge(gothenburg.subdiv, avg_dist_deso, by="deso", all=FALSE)

ggplot() + 
  geom_sf(data = gothenburg.subdiv.merged, aes(fill=avg_dist), size=0.4) +
  scale_fill_distiller(palette = "Oranges")

```

```{r dijkstras_graph_search}
## Alternative

##### TESTING NETWORK MEASURES #####
## The function distances, for example, returns a numeric matrix containing the 
# distances of the shortest paths between every possible combination of nodes. 
# It will automatically choose a suitable algorithm to calculate these shortest paths.

# test.distances <- distances(
#   graph = got_streets_net,
#   weights = got_streets_net %>% activate(edges) %>% pull(length)
# )

####--- Finding the shortest path to the nodes and points (Dijkstra's) (graph search)
#### !!!! WIP This does not work as expected
### Current issues: The complete set of charge points is not being included in these results, and the final path graph is incomplete. 
# From node is the charge point
from_node <- got_streets_net %>%
  activate(nodes) %>%
  #filter(nodeID == 43495) %>% #charge_points.p2p.geo.nodes[1,]$nodeId) %>%
  filter(id == 26205) %>%
  pull(id)

nrow(from_node %>% as_tibble())

non_charge_point <- got_streets_net.merged %>% filter(charge_points == 0) %>% as_tibble() %>% select(id)

# to node is the closest node in the network 
to_node <- got_streets_net %>%
  activate(nodes) %>%
  #filter(nodeID %in% knn.5.nodes$nn.idx[1,]) %>%
  #filter((nodeID %in% c(43494, 43493, 687, 688))) %>%
  filter(id %in% c(2639)) %>%
  pull(id)

nrow(to_node %>% as_tibble())

# path <- sfnetwork(nodes=nodes, edges = edges, directed=FALSE) %>% 
#   st_network_paths(x=., from=from_node, to=to_node)

# Get the shortest path between each of these points
path <- shortest_paths(
  graph = got_streets_net,
  from = from_node,
  to = to_node,
  #mode = "all",
  output="both",
  algorithm ="dijkstra", # Ensure to use Dijkstra's algorithm
  weights = got_streets_net %>% activate(edges) %>% pull(length)
)

path_graph <- got_streets_net %>%
    subgraph.edges(eids = path$epath %>% unlist()) %>%
    as_tbl_graph()

path_graph_df <- path_graph %>% as_data_frame(., what="vertices")

#This subgraph can now be analyzed, for example by calculating the total length of the path
path_graph %>%
  activate(edges) %>%
  as_tibble() %>%
  summarise(length = sum(length))

path_graph %>%
  activate(edges) %>%
  as_tibble() %>%
  arrange(desc(length))

### Validate if the routing algorithm works as expected
charge_points.p2p.geo.nodes %>% filter((nodeId %in% path_graph_df$nodeID)) %>% nrow()

# To use for testing and validation
# knn.5.nodes.filtered$nn.idx[3,]
# closest.nodes.new[3,]

# Missing distances
# row ch_node knn_node distance
# 329 43495   43494    5.652420
# 330 43495   43493    25.250847
# 331 43495     687    0.000000
# 332 43495     688    0.000000
# Can get all distances
# 17  70946   23928    3.107808
# 18  70946   13812    10.546524
# 19  70946   13816    42.985783
# 20  70946   13815    76.786830
df <- got_streets_net %>% activate(nodes) %>% as_tibble()
df %>% filter(nodeID == 687)
joined.nodes %>% as_tibble() %>% mutate(nodeID = as.numeric(nodeID)) %>% arrange(nodeID) %>% select(nodeID, geometry) %>% filter(nodeID == 687)

ggplot() +
  # DeSO Subdivisions
  # geom_sf(data = gothenburg.subdiv, alpha=0.6, linewidth=0.6, colour="#8778D7", fill="#90C59B") + 
  # Road network edges
  geom_sf(data = got_streets_net %>% activate(nodes) %>% as_tibble() %>% st_as_sf(), lwd = 0.1, col = 'grey') +
  # charge points
   geom_sf(data = got_streets_net %>% activate(nodes) %>% as_tibble() %>% filter(id %in% c(from_node, to_node)), lwd = 0.5, col = 'green') +
  # Road network nodes
 # geom_sf(data = joined.nodes %>% filter(id %in% c(from_node, to_node)), size=0.4) +
  # Label nodes
  #geom_sf_label(data= sample.nodes, label=sample.nodes$nodeID) +
  # Dijkstra's paths
   geom_sf(data = path_graph %>% activate(edges) %>% as_tibble() %>% st_as_sf(), lwd = 1, col = 'red')
   # Charge points
   # geom_sf(data = graph.merged %>% activate(nodes) %>% filter(nodeID %in% from_node) %>% as_tibble() %>% st_as_sf(), col = 'yellow', size = 1.8) +
  # # Closest nodes
  #geom_sf(data = got_streets_net.merged %>% activate(nodes) %>% filter(nodeID %in% to_node) %>% as_tibble() %>% st_as_sf(), col = 'blue', size = 1) +
  #geom_sf(data=charge_points.p2p.geo.snaps, col="red") +
  # Coords All paths
  # coord_sf(xlim = c(11.940, 11.946),
  #          ylim = c(57.697, 57.70),
  #          expand = TRUE, clip = "on")
  # # Coords test Missings
  # coord_sf(xlim = c(11.96, 11.9713),
  #          ylim = c(57.706, 57.715),
  #          expand = TRUE, clip = "on")
  # coord_sf(xlim = c(11.71, 12.16),
  #          ylim = c(57.58, 57.86),
  #          expand = TRUE, clip="on")
```

### Accessibility measures

Using a measure for potential accessibility by distance decay. Operationalized as a negative exponential function for surrounding nodes

```{r prepare_data_accessibility, include=FALSE}
# https://cran.r-project.org/web/packages/accessibility/vignettes/accessibility.html
# Setup network to be compatible for accessibility functions

# Setup a distance travel matrix using all nodes and the distance to their knn nearest charge point
distance_travel_matrix <- closest_charge %>% 
  rownames_to_column(., var = "node_id") %>%
  mutate(charge_node_id = mapply(getNodePoint, charge_id)) %>%
  mutate(from_id = node_id, to_id = charge_node_id, 
         length = distance) %>% 
  select(from_id, to_id, length)

# Find  the number of charge points
count_occurance_cp <- closest_charge %>% count(charge_id) %>%
  rownames_to_column(., var = "node_id") %>%
  select(node_id, charge_id, n)

# Create a Land use data frame
charge_land_use <- got_streets_net.merged %>% 
  rename(id = "nodeID") %>% 
  as_data_frame(., what="vertices") %>%
  select(id, charge_points)

```

```{r potential_accessibility_func, include=FALSE}

# Potential function (distance decay)
# The potential accessibility
# measure estimates the accessibility of opportunities in zone i to all other zones (n) in which smaller and/or more distant opportunities provide diminishing influences. The measure has the following form, assuming a negative exponential cost function:
  negative_exp <- gravity(
    distance_travel_matrix,
    charge_land_use,
    opportunity = "charge_points",
    travel_cost = "length",
    decay_function = decay_exponential(decay_value = 0.5)
  )

negative_exp %>% mutate(charge_points = scale(charge_points))

```

### Calculate average distance by DeSO subdivisions

```{r distance_factor_func, include=FALSE}
# Function for creating a level set for distance, to be used for discrete cases
getDistanceLevel <- function(x) {
  if (x > 30) {
    return("30+")
  } else if(x >= 5) {
    if (x < 10) {
      return("5-10")
    } else if (x >= 10 & x < 15) {
      return("10-15")
    } else if (x >= 15 & x < 20) {
      return("15-20")
    } else if  (x >= 20 & x < 25) {
      return("20-25") 
    } else {
      return("25-30")
    }
  } else {
    return("< 5")
  }
}

```

# Spatial Join and merge node and distance averages
```{r merge_nodes_to_desos, include=FALSE}
#####--- Spatial join for GBG DeSO polygons 

nodes.averages <- got_streets_net %>%
  activate(nodes) %>%
  as_tibble() %>% 
  mutate(knn_chargeId = closest_charge$charge_id, 
         knn_distance = closest_charge$distance) 

#Merge new distance decay accessibility values
nodes.averages <- merge(nodes.averages, 
                        negative_exp, 
                        by.x="nodeID", 
                        by.y="id", all=TRUE) %>%
  rename(distance_decay_accessibility = "charge_points") %>%
st_as_sf()

## Filter nodes by subdivision (Spatial join)
filtered.nodes <- st_filter(nodes.averages, gothenburg.subdiv)

joined.nodes <- st_join(filtered.nodes, gothenburg.subdiv, join = st_within, left=TRUE)
joined.nodes <- joined.nodes %>% 
  mutate(knn_distance = knn_distance * 1000,
         distance_decay_accessibility = distance_decay_accessibility * 1000) %>% 
  mutate(distance_level = factor(mapply(getDistanceLevel, knn_distance))) %>%
  mutate(distance_level = fct_relevel(
      distance_level,
      c(
        "< 5",
        "5-10",
        "10-15",
        "15-20",
        "20-15",
        "25-30",
        "30+"
      )
    ))

levels(joined.nodes$distance_level)
avg_dist_deso <- joined.nodes %>% as_tibble() %>% 
  group_by(deso) %>% 
  summarise(nodes = n(),
            avg_dist = mean(knn_distance),
            avg_dd_accessibility = mean(distance_decay_accessibility, na.rm = TRUE)) %>% ### Make sure to handle this na in the results analysis
  arrange(avg_dd_accessibility) ## drop_na() ?

avg_dist_deso <- merge(avg_dist_deso, regSO_deso_table) %>% 
  arrange(desc(avg_dist))

first_geo_by_deso <- joined.nodes %>% as_tibble() %>% 
  group_by(deso) %>%
  filter(row_number()==1) %>%
  rename("deso.ref" = deso) %>%
  select(deso.ref, geometry)

avg_dist_deso_latlong <- merge(x = avg_dist_deso, 
                               y = first_geo_by_deso, 
                               by.x ="deso",
                               by.y = "deso.ref") %>%
  as_tibble()

joined.nodes %>% filter(deso == "1480C3440") %>% nrow()

gothenburg.subdiv <- gothenburg.subdiv %>% as_tibble() %>%
  #filter(deso %in% avg_dist_deso$deso & !is.na(deso)) %>%
  st_as_sf()

#Merge new values to the polygons sf object
gothenburg.subdiv <- merge(gothenburg.subdiv, avg_dist_deso_latlong, by="deso", all=FALSE)

# Create factor levels for distances 
# gothenburg.subdiv <- gothenburg.subdiv %>% mutate(distance_level = factor(mapply(getDistanceLevel, avg_dist)))

# Filter out outliers
#gothenburg.subdiv <- gothenburg.subdiv %>% filter(avg_dist > 50)

```

```{r plot_avg_distance_deso, eval=TRUE, include=TRUE}

### Density plot
gothenburg.subdiv %>% as_tibble() %>% 
  ggplot() + geom_density(aes(avg_dist)) + 
  labs(title="Breakdown distance to public charging points",
       subtitle = "Density",
       fill = "Distance",
       x="Average distance",
       y="") +
  theme_light()

ggplot(gothenburg.subdiv, aes(x=as.factor(regso), y=log(avg_dd_accessibility), size=nodes)) +
  geom_point(alpha=0.7)
# avg_dd_accessibility
gothenburg.subdiv %>% arrange(desc(avg_dd_accessibility))
```

## Analysis

### DeSO subdivision demographics

```{r load_demographic_data}
gothenburg.subdiv <- gothenburg.subdiv %>% arrange(regso)
regso.list <- avg_dist_deso %>% select(deso, regso)
# Read csvs for deso data per demographic
deso.income <- read_csv("data/income_demographics_deso.csv")
deso.foreignborn <- read_csv("data/foreignborn_demographics_deso.csv")
deso.education <- read_csv("data/education_demographics_deso.csv")
deso.housing <- read_csv("data/housing_demographics_deso.csv")
deso.cars <- read_csv("data/cars_demographics_deso.csv")
deso.age <- read_csv("data/age_demographics_deso.csv")
deso.households <- read_csv("data/households_demographics_deso.csv")
```

```{r clean_demographic_data}
# Cleaning
deso.foreignborn <- deso.foreignborn %>% 
  mutate(swedish = swedish / total, foreign = foreign / total) %>% 
  select(-total) %>%
  filter(deso %in% regso.list$deso)
deso.income <- deso.income %>% 
  select(deso, `Median income tkr`) %>%
  filter(deso %in% regso.list$deso)
deso.education <- deso.education %>% 
  filter(deso %in% regso.list$deso) %>%
  mutate(total = rowSums(.[, -1]),
         middle_school = middle_school / total,
         gymnasium = gymnasium / total,
         post_gymnasium_less_than_3 = post_gymnasium_less_than_3 / total,
         post_gymnasium_more_than_3 = post_gymnasium_more_than_3 / total,
         education_missing = education_missing / total) %>%
  select(-total)
deso.housing <- deso.housing %>% 
  filter(deso %in% regso.list$deso) %>%
  select(deso, hyresrätt, bostadsrätt, äganderätt) %>%
  filter(deso %in% regso.list$deso) %>%
  mutate(total = rowSums(.[, -1]),
         hyresrätt = hyresrätt / total,
         bostadsrätt = bostadsrätt / total,
         äganderätt = äganderätt / total) %>%
  select(-total)
deso.cars <- deso.cars %>%
  filter(deso %in% regso.list$deso) %>% 
  select(deso, bilar_i_trafik)

deso.age <- deso.age %>% 
  filter(deso %in% regso.list$deso) %>%
  # Create percentage of ages
  mutate(across(colnames(deso.age[,3:19]), function(x) x/total))

deso.pop <- deso.age %>% 
  filter(deso %in% regso.list$deso) %>% 
  mutate(pop = total, 
         ages = list(data_frame(across(colnames(deso.age[,3:19]))))) %>% 
  select(deso, pop, ages)

deso.households <- deso.households %>%
  filter(deso %in% regso.list$deso) %>% 
  select(deso, total_households) 
deso_aream2 <- gothenburg.subdiv %>% as_tibble() %>% 
  mutate(area_m2 = as.double(area)) %>% 
  select(deso, area_m2)
```

```{r combine_demographic_data}
# Combine demographic data with distances by DeSO subdivision
deso.demographics <- deso_aream2 %>%
  merge(., deso.income) %>%
  merge(., avg_dist_deso, by="deso") %>%
  merge(., deso.pop, by="deso") %>%
  merge(., deso.households, by="deso") %>%
  merge(., deso.foreignborn, by="deso") %>%
  merge(., deso.education, by="deso") %>%
  merge(., deso.housing, by="deso") %>%
  merge(., deso.cars, by="deso") %>%
  # Calculate average cars per household
  mutate(cars_per_household = bilar_i_trafik / total_households) %>%
  select(deso, regso, nodes, avg_dist,  everything()) %>%
  arrange(desc(avg_dist))

view(deso.demographics.rentals)

deso.demographics.WOMajorityHouses <- deso.demographics %>%
  # Filter out desos with majority houses
  filter(äganderätt < 0.5)

deso.demographics.rentals <- deso.demographics %>%
  # Filter out desos with majority houses
  filter(hyresrätt > 0.50)

deso.demographics.owner.occupied <- deso.demographics %>%
filter(äganderätt >= 0.75 | bostadsrätt >= 0.75)

# deso.demographics.rentals %>% arrange(desc(avg_dist)) %>% view()
# avg_dist_deso %>% arrange(desc(avg_dist)) %>% view()
regso.list %>% arrange(regso)

# # Write files to csv if needed
# write_csv(deso.demographics %>% select(-ages), "data/deso_demographics.csv")
# write.csv(deso.demographics.rentals %>% select(-ages), "data/deso_demographics_rentals.csv")
```

### Logistic regressions

```{r lm_models_all, eval=TRUE, include=TRUE}
model1 <- lm(avg_dist~foreign+nodes+äganderätt+total_households, 
             data = deso.demographics)
summary(model1)
model2 <- lm(avg_dist~foreign+nodes+äganderätt+`Median income tkr`,
             data = deso.demographics)
summary(model2)
model3 <- lm(avg_dist~foreign+nodes+äganderätt+post_gymnasium_more_than_3, 
             data = deso.demographics)
summary(model3)
```

```{r lm_models_rental_areas, eval=TRUE, include=TRUE}
model1.rentals <- lm(avg_dist~foreign+nodes+total_households, 
             data = deso.demographics.rentals)
summary(model1.rentals)
model2.rentals <- lm(avg_dist~foreign+nodes+`Median income tkr`, 
             data = deso.demographics.rentals)
summary(model2.rentals)
model3.rentals <- lm(avg_dist~foreign+nodes+äganderätt+post_gymnasium_more_than_3, 
             data = deso.demographics.rentals)
summary(model3.rentals)
```

```{r plot_demographic_results, eval=TRUE, include=TRUE}
#scale(deso.demographics$avg_dist)
ggplot(data = deso.demographics, aes(x=avg_dist, y=total_households)) + 
  geom_point(aes(color=ifelse(foreign > 0.49, "foreign majority", "swedish majority"))) +
  labs(color="Swedish vs foreign")
  #scale_x_continuous(limits = c(0, 40))
```

```{r plot_demographic_results.rentals, eval=TRUE, include=TRUE}
# Plot rental-majority DeSOs based on dijkstras shortest distance to charge point
ggplot() + 
  geom_sf(data = gothenburg.subdiv.merged %>% filter(deso %in% deso.demographics.rentals$deso), aes(fill=avg_dist), size=0.4) +
  scale_fill_distiller(palette = "Oranges")
```

### PCA analysis

<https://www.datacamp.com/tutorial/pca-analysis-r>

#### All subdivisions

```{r setup_pca_matrix.all}
#### PCA and Correlations matrix
deso.demographics.pca <- deso.demographics %>% select(-deso, -regso, -nodes, -ages)
deso.demographics.normalized <- scale(deso.demographics.pca)
corr_matrix <- cor(deso.demographics.normalized)
```

#### Majority rental subdivisions

```{r setup_pca_matrix.rentals}
#### PCA and Correlations matrix
deso.demographics.pca <- deso.demographics.rentals %>% select(-deso, -regso, -nodes, -ages)
deso.demographics.normalized <- scale(deso.demographics.pca)
corr_matrix <- cor(deso.demographics.normalized)
```

```{r setup_pca_matrix.owner_occupied}
#### PCA and Correlations matrix
deso.demographics.pca <- deso.demographics.owner.occupied %>% select(-deso, -regso, -nodes, -ages)
deso.demographics.normalized <- scale(deso.demographics.pca)
corr_matrix <- cor(deso.demographics.normalized)
```
#### Correlation Matrix

```{r correlation_plot}
# Plot correlation matrix
ggcorrplot(corr_matrix, 
           title = "Correlation matrix, GOT DeSO demographic data",
           )
```

```{r plot_pca_eigan_values}
# Perform PCA and run analysis on variables and relations
data.pca <- princomp(corr_matrix)
summary(data.pca)
data.pca$loadings[, 1:2]
fviz_eig(data.pca, addlabels = TRUE)
fviz_pca_var(data.pca, col.var = "black")
fviz_cos2(data.pca, choice = "var", axes = 1:2)
fviz_pca_var(data.pca, col.var = "cos2",
             gradient.cols = c("black", "orange", "green"),
             repel = TRUE)
```
